---
title: "Reproducible Research: Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the .csv file.

```{r}
data <- read.csv("activity.csv", colClasses = c("numeric", "character", "integer"))
```

Format date and group data by day using group_by function in dyplr

```{r}
library(dplyr)
totsteps <- data %>%
filter(!is.na(steps)) %>%
group_by(date) %>%
summarize(steps = sum(steps))
```

Plot histogram of total steps per day

```{r}
hist(totsteps$steps, breaks=20, xlab='steps per day' )
dev.copy(png, file="plot_step2.png", width=480, height=480)
dev.off()

```

Calculate mean and median steps

```{r}
median(totsteps$steps)
mean(totsteps$steps)
```

Group data by interval and calcaulate average number of steps

```{r}
int <- data %>%
filter(!is.na(steps)) %>%
group_by(interval) %>%
summarize(steps = mean(steps))
```

Plot steps taken as function of interval

```{r}
plot(int)
dev.copy(png, file="plot_step4.png", width=480, height=480)
dev.off()

```

Find the interval where the maximum steps are taken, on average

```{r}
int[which.max(int$steps),]
```

Replace missing values with average number of steps per interval, and print total number of missing values

```{r}
alldata <- data
miss <- is.na(alldata$steps)
print(sum(miss))
avgval <- tapply(alldata$steps, alldata$interval, mean, na.rm=TRUE, simplify=TRUE)
alldata$steps[miss] <- avgval[as.character(alldata$interval[miss])]
```

Group by date again, and plot new histogram with imputed data included

```{r}
totsteps2 <- alldata %>%
filter(!is.na(steps)) %>%
group_by(date) %>%
summarize(steps = sum(steps))

hist(totsteps2$steps, breaks=20, xlab='steps per day (missing data replaced with average)')
dev.copy(png, file="plot_step7.png", width=480, height=480)
dev.off()

```

Select data from Saturday, Sunday

```{r}
library(lubridate)
alldata <- mutate(alldata, weekendorday = ifelse(weekdays(ymd(alldata$date)) == "Saturday" | weekdays(ymd(alldata$date)) == "Sunday", "Weekends", "Weekdays"))
alldata$weekendorday <- as.factor(alldata$weekendorday)
```

```{r}
interval_full <- alldata %>% group_by(interval, weekendorday) %>% summarise(steps = mean(steps))
library(ggplot2)
pp <- ggplot(interval_full, aes(x=interval, y=steps, color = weekendorday)) +
  geom_line() +
  facet_wrap(~weekendorday, ncol = 2, nrow=1)
print(pp)
dev.copy(png, file="plot_step8.png", width=480, height=480)
dev.off()

```